---
title: "Quick start: US Presidential Inaugural Addresses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick start: US Presidential Inaugural Addresses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 6, dpi = 72,
  cache = TRUE
)
```

This vignette reproduces the figures used in Section 4 of the
companion paper, applied to the US Presidential Inaugural
Addresses corpus shipped with the package.

```{r setup}
library(ljmds)
```

## 1. Load the corpus from CSV

The corpus ships as a CSV in which column 1 is the calendar year
and columns 2.. are 0/1 indicators of keyword presence (no source
text is included).

```{r load}
d <- ljmds.read.csv("inaugural")
dim(d$X)                  # 59 x 106
range(d$t)                # 1789 2021
head(d$keywords, 10)
```

## 2. Joint (h, k) selection

A minimal grid is used here so the vignette builds quickly;
a finer grid is recommended in practice (e.g.,
`h.grid = c(8, 10, 15, 20, 25, 30, 40, 50, 70, 100)`).
The trivial `k = 2` split is excluded by leaving 2 out of
`k.grid` (default `3:6`).

```{r select}
sel <- ljmds.select(d$X, d$t,
                    h.grid = c(30, 50),
                    k.grid = 3:4)
sel$h.hat
sel$k.hat
sel$S.hat
round(sel$S, 3)
```

## 3. Run the pipeline at $(k, h) = (3, 50)$

```{r fit}
fit <- ljmds.pipeline(d$X, d$t, h = 50, k = 3)
```

## 4. Figures from the paper

### Figure 3: cluster centroid trajectories

```{r fig3, fig.cap = "Centroid trajectories on the modified MDS configuration."}
plot(fit, type = "trajectory")
```

### Figure 4: Ward dendrogram

```{r fig4, fig.cap = "Ward dendrogram on the trajectory distance H."}
plot(fit, type = "dendrogram")
```

### Figure 5: joint selection heatmap

```{r fig5, fig.cap = "Silhouette heatmap S(k, h) with maximizer marker."}
plot(sel)
```

### Figure 6: time-collapsed MDS map

```{r fig6, fig.cap = "Classical MDS of H, coloured by class."}
plot(fit, type = "cmd")
```

### Figure 7: class mean curves

```{r fig7, fig.cap = "Class mean occurrence curves."}
plot(fit, type = "means")
```

### Figure 8: per-class small multiples

```{r fig8, fig.cap = "Individual smoothed curves and class means."}
plot(fit, type = "panels")
```

### Figure 9: animated trajectory map (GIF)

A pre-rendered animation ships with the package; locate it
with `system.file()` and view it directly:

```{r gif-path}
gif_path <- system.file("extdata", "inaugural.gif",
                        package = "ljmds")
gif_path
```

```{r view-gif, eval = FALSE}
browseURL(gif_path)              # open in default browser
# magick::image_read(gif_path)   # or: open in RStudio Viewer
```

```{r show-gif, echo = FALSE, out.width = "100%"}
knitr::include_graphics(gif_path)
```

To regenerate the GIF from the fitted object (writes a new
file to the current working directory; takes about a minute),
uncomment and run:

```{r regen-gif, eval = FALSE}
# gif <- ljmds.animate(fit, file = "inaugural.gif",
#                   trail = 7, fps = 2)
```
