---
title: "Quick start: Peace Declaration of Hiroshima"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick start: Peace Declaration of Hiroshima}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 6, dpi = 72,
  cache = TRUE
)
```

This vignette reproduces the figures used in Section 5 of the
companion paper, applied to the Peace Declaration of Hiroshima
1947--2025 (the 1950 declaration was not delivered).

```{r setup}
library(ljmds)
```

## 1. Load the corpus from CSV

```{r load}
d <- ljmds.read.csv("peace_declaration")
dim(d$X)                  # 78 x 95
range(d$t)                # 1947 2025
head(d$keywords, 10)
```

## 2. Joint (h, k) selection

A minimal grid is used here so the vignette builds quickly;
a finer grid is recommended in practice (e.g.,
`h.grid = c(3, 4, 5, 6, 8, 10, 12, 15, 20)`).  The trivial
`k = 2` split is excluded by leaving 2 out of `k.grid`.

```{r select}
sel <- ljmds.select(d$X, d$t,
                    h.grid = c(8, 12),
                    k.grid = 3:4)
sel$h.hat
sel$k.hat
sel$S.hat
round(sel$S, 3)
```

## 3. Run the pipeline at $(k, h) = (4, 8)$

```{r fit}
fit <- ljmds.pipeline(d$X, d$t, h = 8, k = 4)
```

## 4. Figures from the paper

```{r fig10, fig.cap = "Cluster centroid trajectories."}
plot(fit, type = "trajectory")
```

```{r fig11, fig.cap = "Ward dendrogram on H."}
plot(fit, type = "dendrogram")
```

```{r fig12, fig.cap = "Silhouette heatmap with maximizer marker."}
plot(sel)
```

```{r fig13, fig.cap = "Time-collapsed MDS map."}
plot(fit, type = "cmd")
```

```{r fig14, fig.cap = "Class mean occurrence curves."}
plot(fit, type = "means")
```

```{r fig15, fig.cap = "Per-class smoothed curves."}
plot(fit, type = "panels")
```

A pre-rendered animation ships with the package; locate it
with `system.file()` and view it directly:

```{r gif-path}
gif_path <- system.file("extdata", "peace_declaration.gif",
                        package = "ljmds")
gif_path
```

```{r view-gif, eval = FALSE}
browseURL(gif_path)              # open in default browser
# magick::image_read(gif_path)   # or: open in RStudio Viewer
```

```{r show-gif, echo = FALSE, out.width = "100%"}
knitr::include_graphics(gif_path)
```

To regenerate the GIF from the fitted object (writes a new
file to the current working directory; takes about two
minutes), uncomment and run:

```{r regen-gif, eval = FALSE}
# gif <- ljmds.animate(fit, file = "peace_declaration.gif",
#                   trail = 10, fps = 2)
```
